---
title: "Creating Static and Interactive Maps in R"
author: "Cian White"
date: '2020-03-26'
slug: Mapping_in_R
categories:
  - R
tags:
  - Mapping
  - R Markdown
output:
  blogdown::html_page:
    toc: true
---

```{r echo=FALSE}
bytes <- file.size("Mapping_Script.Rmd")
words <- bytes/10
minutes <- words/200
```
Reading time: 13 minute(s) @ 200 WPM.

*A note: if you are reading this on Rpubs, the links will only work is you `Control` Click.*

# Irish COVID 19 Intertactive Map
***

### Aim
To create a map that tracks the spread of COVID 19 in Ireland as the pandemic progresses. It is meant as a public educational tool and data will be collected from various sources, such as those cited in the Wiki page on COVID 19 in Ireland, and also health statistics which the HSE have recentrly started to publish on a county by county level.

The code was inspired by Dr. [Richard Lent's notebook](https://richardlent.github.io/rnotebooks/maps.nb.html) on how to create interactive maps in R.

The book [Geocomputation with R](https://geocompr.robinlovelace.net/) is a very useful resource. Excellent to get you started creating maps in R. Covers all the spatial data processing that occurs here, reading in shapefiles, manipulating spatial data and adding ancillary data that will be mapped to the spatial data files.

Most of the data handling carried out here uses the [tidyverse](https://www.tidyverse.org/) syntax. The book [R for Data Science](https://r4ds.had.co.nz/) by [Hadley Wickham](http://hadley.nz/) is the best reference source to get you started on using the tidyverse packages. I highly recommend the book.



## Shapefiles
***

For manipulating geometry data see [here](https://geocompr.robinlovelace.net/spatial-class.html)

Using the `sf` package, standing for simple features. It's a widely compatible spatial data format and can be used to read in and manipulate shapefiles.

I downloaded shapefiles for the Republic of Ireland from the [Central Statistics Office](https://www.cso.ie/en/census/census2011boundaryfiles/), specifically the 	`Census2011_Admin_Counties_generalised20m.zip`. I unzipped to the project I am working at and stored it at this location: `Shapefiles/Counties/ROI_Counties/`. I downloaded the Northern Ireland county shapefile from the Northern Ireland [Open Data Initiative](https://www.opendatani.gov.uk/about). The shapefile can be found [here](https://www.opendatani.gov.uk/dataset/osni-open-data-50k-boundaries-ni-counties)

Creating all ireland shapefile.
```{r message=FALSE}
library(dplyr)
library(tidyverse)
library(sf) #for reading in shapefiles
```

```{r}
ROI_counties = st_read("Shapefiles/Counties/ROI_Counties/Census2011_Admin_Counties_generalised20m.shp")

NI_counties = st_read("Shapefiles/Counties/NI_Counties/OSNI_Open_Data_50K_Admin_Boundaries_–_Counties.shp")


ROI_counties <- select(ROI_counties, c(Nation = NUTS1NAME, County = COUNTYNAME, pop = TOTAL2011, ID = COUNTY))

NI <- NI_counties %>%
          mutate(Nation = "Northern Ireland", Province = "Ulster",
          pop = c(618108, 174792, 531665, 61170, 247132, 179000),
          Full_County = c("Antrim", "Armagh", "Down", "Fermanagh",                                "Derry", "Tyrone"),
          County_ID =c(27:32)) %>%
  select(Nation, Full_County, County_ID, Province, pop)

county <- c("Carlow", "Dublin", "Dublin", "Dublin", "Dublin", "Kildare", "Kilkenny", "Laois", "Longford", "Louth", "Meath", "Offaly", "Westmeath", "Wexford", "Wicklow", "Clare", "Cork", "Cork", 
            "Kerry", "Limerick", "Limerick", "Tipperary", "Tipperary", "Waterford", "Waterford", "Galway", "Galway", "Leitrum", "Mayo", "Roscommon", "Sligo", "Cavan", "Donegal",                       "Monaghan")

provinces <- c(rep("Leinster",times = 15), rep("Munster", times = 10),
               rep("Connacht", times =6), rep("Ulster", times =3))

County_ID <- c(1, rep(2, times=4), 3:13, rep(14,times=2), 15, rep(16, times=2), rep(17, times=2), rep(18,times=2), rep(19, times=2), 20:26)

ROI <- ROI_counties %>%
          arrange(ID) %>%
          mutate(Province = provinces, Full_County = county,
                 County_ID = County_ID) %>%
  select(Nation, Full_County, County_ID, Province, pop)


Ireland <- rbind(ROI,NI)

Ireland$Full_County <- as.factor(Ireland$Full_County)

names(Ireland) <-c("Nation", "County", "County_ID", "Province", "pop", "geometry")
# Add fill and border layers to nz shape
```



## Covid data
***

I am reading in the Covid data from dataset I created from the data released by the [Department of Health](https://www.gov.ie/en/news/7e0924-latest-updates-on-covid-19-coronavirus/) website.

A note: I am creating multiple spatial datasets here using the `group_by` and `summarise` functions from `dplyr`. The `sp` package handles spatial data manipulation, and is compatible with `dplyr`. Each time I `group_by`and `summarise` I am creating a new dataset, with aggregated geometries. For example, here I group by county, province, nation and island, and can create plots at each level of grouping. For more on this see [here](https://geocompr.robinlovelace.net/geometric-operations.html#fig:us-regions)

Reading in county, province and national data and joining to shapefiles so it can be mapped.
```{r message="center"}
#reading in health data and tidying up
County_data <- as_tibble(read.csv("Data/County.csv")) %>%
  pivot_longer(cols = starts_with("x"), names_to = "Date",
               values_to = "Numbers") %>%
  mutate(Type = "Cases") %>%
  pivot_wider(names_from = Type, values_from = Numbers)

#install.packages("lubridate")#for manipulating dates and times
library(lubridate)

County_data$Date <- str_replace(County_data$Date, "^[A-Z]*", "") %>%
  dmy()

County_data$Cases <- as.numeric(str_replace(County_data$Cases, "<", ""))

#setting column names so that `County` column matches in Ireland_projected and County data to join
names(County_data) <- c("County", "Province", "Nation", "Date", "Cases")

#Grouping by County level in shapefile
County_Level = Ireland %>%
  group_by(County) %>%
  summarize(pop = sum(pop, na.rm = TRUE),
            Nation = first(Nation),
            Province = first(Province))

#joining data
County_stats <- left_join(County_Level,County_data)
County_stats <- County_stats %>%
  mutate(rate = Cases/pop)




#Province Data: grouping by province to and date to get province level data per date, summing over cases
Province_data = County_data %>%
  group_by(Province, Date) %>%
  summarize(Cases = sum(Cases, na.rm = TRUE))

#as ulster is split by nation, bringing in Northern Ireland data
#Loading Nation, to read in ulster values for Province data
Nation <- as_tibble(read.csv("Data/Nation.csv")) %>%
  pivot_longer(cols = starts_with("x"), names_to = "Date",
               values_to = "Numbers") %>%
  pivot_wider(names_from = Type, values_from = Numbers)

Nation$Date <- str_replace(Nation$Date, "^[A-Z]*", "") %>%
  dmy()

SixCounties <- filter(Nation, Nation == "Northern Ireland",
                      Date %in% unique(Province_data$Date))
#adding ulster data to province dataset
Province_data$Cases[Province_data$Province == "Ulster"] <- Province_data$Cases[Province_data$Province == "Ulster"] + SixCounties$Cases


#Grouping by Province Level shapefile
Province_Level = Ireland %>%
  group_by(Province) %>%
  summarize(pop = sum(pop, na.rm = TRUE))

#joining province COVID data to geometry data
Province_stats <- left_join(Province_Level, Province_data)
Province_stats <- Province_stats %>%
  mutate(rate = Cases/pop)




#National Data
#importing national level health data
Nation <- as_tibble(read.csv("Data/Nation.csv")) %>%
  pivot_longer(cols = starts_with("x"), names_to = "Date",
               values_to = "Numbers") %>%
  pivot_wider(names_from = Type, values_from = Numbers)

#editing date columns
Nation$Date <- str_replace(Nation$Date, "^[A-Z]*", "") %>%
  dmy()

#Grouping by Nation in shapefile and summing populations in each nation.
National_agg <- Ireland %>%
  group_by(Nation) %>%
  summarize(pop = sum(pop, na.rm = TRUE))

#joining province COVID data to geometry data
National_stats <- left_join(National_agg, Nation)
```


## Using ggplot to make maps
***

`ggplot`is can plot maps using the `geom_sf` function. All the functionality of `ggplot` can be used. While static maps are created, these can be made interactive by using the `plotly`package - see [here](https://plotly-r.com/overview.html) for the book on `plotly`.

It is also possible to animate ggplots using the `gganimate` package, creating a `.gif` that is a composite of many static map images. The [Gapminder](https://www.gapminder.org/tools/#$chart-type=bubbles) website uses this software to generate it's amazing images. For more info see the [gganimate](https://gganimate.com/) webpage.

Here, I am just creating a static map of Ireland, with counties coloured by the number of COVID 19 cases.

```{r message=FALSE, fig.align="center"}
library(ggplot2)

g1 = ggplot(County_stats) +
  geom_sf(aes(fill = Cases))
 
g1
```


## Creating Thematic Maps
***

The `tmap` package is a powerful and flexible map-making package with sensible defaults. It has a concise syntax that allows for the creation of attractive maps with minimal code which will be familiar to `ggplot2` users. It also has the unique capability to generate static and interactive maps using the same code via `tmap_mode()`.

For more see [here](https://geocompr.robinlovelace.net/adv-map.html).

```{r message=FALSE, fig.align="center"}
library(tmap)
tmap_mode("plot")

Ireland %>%
  mutate(pop_100 = round(pop/1000)) %>%
tm_shape() +
  tm_fill(col = "pop_100",
          title = "Population (in 1000s)",
          text.size = 0.5,
          position = c("top", "left")) +
  tm_scale_bar(breaks = c(0, 50, 100), text.size = 0.5,
               position = c("right", "bottom")) +
  tm_layout(frame = FALSE)


# County Map
 Ireland %>%
  group_by(County) %>%
  summarize(pop = sum(pop, na.rm = TRUE)) %>%
  tm_shape() +
  tm_fill(col = "pop",
          aplha = 0.5,
          title = "Population (in 1000s)",
          text.size = 0.5,
          position = c("top", "left")) +
  tm_scale_bar(breaks = c(0, 50, 100), text.size = 0.5,
               position = c("right", "bottom")) +
  tm_layout(frame = FALSE)

# Province Map
 Ireland %>%
  group_by(Province) %>%
  summarize(pop = sum(pop, na.rm = TRUE)) %>%
  tm_shape() +
  tm_fill(col = "pop",
          aplha = 0.5,
          title = "Population (in 1000s)",
          text.size = 0.5,
          position = c("top", "left")) +
  tm_scale_bar(breaks = c(0, 50, 100), text.size = 0.5,
               position = c("right", "bottom")) +
  tm_layout(frame = FALSE)

 
# Nation Map
  Ireland %>%
  group_by(Nation) %>%
  summarize(pop = sum(pop, na.rm = TRUE)) %>%
  tm_shape() +
  tm_fill(col = "pop",
          aplha = 0.5,
          title = "Population (in 1000s)",
          text.size = 0.5,
          position = c("top", "left")) +
  tm_scale_bar(breaks = c(0, 50, 100), text.size = 0.5,
               position = c("right", "bottom")) +
  tm_layout(frame = FALSE)

tmap_mode("view")

  Ireland %>%
  mutate(pop_100 = round(pop/1000)) %>%
tm_shape() +
  tm_fill(col = "pop_100",
          title = "Population (in 1000s)",
          text.size = 0.5,
          position = c("top", "left")) +
  tm_scale_bar(breaks = c(0, 50, 100), text.size = 0.5,
               position = c("right", "bottom")) +
  tm_layout(frame = FALSE)
```


## Animated Maps
***
Just like `gganimate`, it is possible to make `.gif` files using the `tmap` package and another package called `magick`.

An example of a `.gif` created this way can be found [here](https://geocompr.robinlovelace.net/adv-map.html#fig:animus).

You will need to download Image Magick from  [here](https://imagemagick.org/script/download.php).

During installation you must check yes to downloading the `converter` function. Be careful to do so or the process won't work.

Finally, you need to include the Image Magick application in your PATH. The code `Sys.setenv(PATH = paste("C:\\Program Files\\ImageMagick-7.0.10-Q16", Sys.getenv("PATH"), sep=";"))` adds Image Magick to the PATH, but you will need to check where Image Magick is stored on your computer and add the file address, replacing `C:\\Program Files\\ImageMagick-7.0.10-Q16`. On a windows system make sure to include two back slashes, rather than one, in the address.

```{r message=FALSE, animation.hook = 'ffmpeg'}
#creating an animated Gif using tmap and magick packages
library(magick)
  
Nation_anim = tm_shape(National_stats) + 
    tm_fill(col = "Cases",
          aplha = 0.5,
          title = "Cases",
          text.size = 0.5,
          position = c("top", "left")) +
  tm_scale_bar(breaks = c(0, 50, 100), text.size = 0.5,
               position = c("right", "bottom")) +
  tm_layout(frame = FALSE) +
  tm_facets(along = "Date", free.coords = FALSE)

#adding image magick to file path. 
Sys.setenv(PATH = paste("C:\\Program Files\\ImageMagick-7.0.10-Q16", Sys.getenv("PATH"), sep=";"))
#tmap::tmap_animation(tm =Nation_anim, filename = "Nation_anim.gif", delay = 25, width = 1200, height = 800)
magick::image_read("Nation_anim.gif")
```

As you can see from above, a `.gif` is really just a series of static images shown at a certain interval. For some reason, the `.gif` isn't loading into the markdown document, but you can see the gif at my [github page](https://github.com/ciwhite/Mapping/blob/master/Nation_anim.gif).

## Using Leaflet
***

Last but not least is `leaflet`. In the developers' own words:

> an open-source JavaScript library
for mobile-friendly interactive maps

It's used by the Washington Post, Facebook, Finanicial Times to create beautiful interactive maps that engage the public.

And it has an R package associated with it.

It is the most mature and widely used interactive mapping package in R. leaflet provides a relatively low-level interface to the Leaflet JavaScript library and many of its arguments can be understood by reading the documentation of the original JavaScript library (see [leafletjs.com](https://leafletjs.com/)).

Leaflet maps are created with leaflet(), the result of which is a leaflet map object which can be piped to other leaflet functions. This allows multiple map layers and control settings to be added interactively

For more on leaflet in R see the [GitHub page](https://rstudio.github.io/leaflet/)

Let's create a basic interactive map first of all.
```{r fig.align="center", message=FALSE}
library(RColorBrewer)
library(leaflet)
library(widgetframe)

County_stats = st_transform(County_stats, 4326) #projecting County_stats spatial data onto WGS84 for leaflet

#setting columns names
#names(Ireland_projected) <-c("Nation", "County", "County_ID", "Province", "pop", "geometry")


pal <- colorNumeric(
    palette = "Reds",
    domain = County_stats$Cases)

highlight <- highlightOptions(color = "grey", 
                              weight = 2,
                              bringToFront =TRUE)

County_map <- leaflet() %>%
  #addProviderTiles(providers$Esri.WorldTopoMap)%>% #can add a background map by uncommenting
    addPolygons(data = County_stats, 
                fillColor = ~pal(Cases), 
                color = "#b2aeae", # Need to use hex color codes.
                fillOpacity = 0.7, 
                weight = 0.5, 
                smoothFactor = 0.2,
                highlightOptions = highlight) %>%
  addLegend(pal = pal,
              values = 
                County_stats$Cases,
              position = "bottomright", 
              title = "Cases<br>") %>%
    addScaleBar(position = "topright")
    
# Display and save the map
print(County_map)

```



Ramping up the settings
```{r fig.align="center", message=FALSE, out.width = "100%"}
library(scales)# inserts comma's into large numbers for use on labels through comma()

#setting up label
countypolylabel <- paste0("<strong>", County_stats$County,"</strong>", "<br>", "Population: ", comma(round(County_stats$pop,-3)))

countycirlelabel <- paste0("<strong>","Cases: ","</strong>",
                     County_stats$Cases[County_stats$Date == "2020-03-20"],"<br>",
                     "<strong>", "Infection Rate: ","</strong>",
                     round(County_stats$rate[County_stats$Date == "2020-03-20"],6))

#getting centroids for each county so I can locate a circle in each county
county_centroid = st_centroid(County_stats)

countylng <- vector("double", 32)
countylat <- vector("double", 32)

for (i in seq_along(county_centroid$geometry)){
countylng[i] <- c(county_centroid$geometry[[i]][1])
countylat[i] <- c(county_centroid$geometry[[i]][2])
}

#setting label options
labeloptions <- labelOptions(style = list("font-weight" = "normal",
                               padding = "3px 8px"),
                             textsize = "15px",
                             direction = "auto")


#colour palette
countypalrate <- colorNumeric(
    palette = "Reds",
    domain = County_stats$rate)

countypalcase <- colorNumeric(
    palette = "Reds",
    domain = County_stats$Cases)

#setting highlight options
highlight <- highlightOptions(color = "grey",weight = 2,bringToFront =TRUE)


#creating map
County_map <- leaflet() %>%
  addMapPane("background_map", zIndex = 410) %>%  # Level 1: bottom
  addMapPane("polygons", zIndex = 420) %>%        # Level 2: middle
  addMapPane("circles", zIndex = 430) %>%          # Level 3: top
  addProviderTiles(providers$Esri.WorldTopoMap,
                   options = pathOptions(pane = "background_map"))%>%
    addPolygons(data = County_stats,# adding a polygon
                fillColor = ~countypalrate(rate),#coluring by rate of infection
                color = "#b2aeae", # Need to use hex color codes.
                fillOpacity = 0.7, 
                weight = 0.5, 
                smoothFactor = 0.2,
                label = lapply(countypolylabel, htmltools::HTML),
                labelOptions = labeloptions,
                highlightOptions = highlight,
                options = pathOptions(pane = "polygons")) %>% #adding to polygon level
  addCircles(data = County_stats,#adding circles
               lng = countylng[seq(1, length(countylng), length(unique(na.omit(County_stats$Date))))], #at latitude of county centroid
               lat = countylat[seq(1, length(countylat), length(unique(na.omit(County_stats$Date))))],
               fillColor = ~countypalcase(Cases), #at longitude of county centroid
               fillOpacity = 0.3,
               weight = 0.5, 
               color = "#FFFFFF",
               radius = 600*sqrt(County_stats$Cases)
               [County_stats$Date == "2020-03-20"], # size of circle
               label = lapply(countycirlelabel, htmltools::HTML),
               labelOptions = labeloptions,
               highlightOptions = highlight,
               options = pathOptions(pane = "circles")) %>% #adding to cicrle level.
    addLegend(pal = countypalrate,
              values = 
                County_stats$rate[County_stats$Date == "2020-03-20"],
              position = "bottomright", 
              title = "Infection Rate<br>") %>%
    addScaleBar(position = "topright")
    
# Display and save the map
invisible(print(County_map))
```

Adding layers to a map using groups
```{r fig.align="center", out.width = "100%"}
#run tranformation when first running chunk
Province_stats = st_transform(Province_stats, 4326) 

##setting up labels 
Province_labels = vector()
#for loop to create labels
for(i in seq_along(unique(Province_stats$Province))){
for(e in seq_along(unique(Province_stats$Date))){
label <- paste0("<strong>",
                      Province_stats$Province[Province_stats$Date ==
                                                unique(Province_stats$Date)[e] &
                        Province_stats$Province ==
                                                unique(Province_stats$Province)[i]],
                      "</strong>", "<br>", "Cases: ",
                      Province_stats$Cases[Province_stats$Date ==
                                                unique(Province_stats$Date)[e] &
                        Province_stats$Province ==
                                                unique(Province_stats$Province)[i]])
Province_labels[[length(Province_labels)+1]] = label
}}


#subsetting labels for by date so I can plot each date as a layer
#using seq() funciton combined with nth() function to pick out each labels for each date.
#Attempting to automate for when ever a new date is added using for loop

Province_labs <- vector()
for (i in seq_along(unique(Province_stats$Date))){
labs <- Province_labels[seq(nth(seq_along(unique(Province_stats$Date)),i),last(seq_along(Province_stats$Date)), by = length(unique(Province_stats$Date)))]
Province_labs <- rbind(Province_labs,labs)
}
#now province_labs contains labels for each dates for all provinces in a row, so can access using square bracket and the order of the unique date.
Province_labs[2,]

unique_date <- unique(Province_stats$Date)

provcirlelabel <- paste0("<strong>","Infection Rate: ","</strong>",
                     round(Province_stats$rate[
                       Province_stats$Date ==
                         last(unique(Province_stats$Date))],5))


# getting centroids for each county for plotting the circles
province_centroid = st_centroid(Province_stats[Province_stats$Date == last(unique(Province_stats$Date)),])

#extracting the longitude and latitude of the centroids to centre the circles at.
provlng <- vector("double", length(unique(Province_stats$Province))*length(unique(Province_stats$Date)))
provlat <- vector("double", length(unique(Province_stats$Province))*length(unique(Province_stats$Date)))

for (i in seq_along(province_centroid$geometry)){
provlng[i] <- c(province_centroid$geometry[[i]][1])
provlat[i] <- c(province_centroid$geometry[[i]][2])
}

#colour palettes
#palette for circles
provpalrate <- colorNumeric(
    palette = "Purples",
    domain = c(0:max(Province_stats$rate)))
#palette for polygons
provpalcases <- colorNumeric(
    palette = "Reds",
    domain = c(0:max(Province_stats$Cases)))

#highlight options for when hovering over with mouse
highlight <- highlightOptions(color = "grey",weight = 2,bringToFront =TRUE)



# getting centroids for each county for plotting the circles
names(National_stats) <- c("Nation", "pop", "geometry", "Island", "Date", "Cases", "Death", "Recovered")

# getting centroids of island of Ireland so can specify the map to always generate at those centroids
National_stats = st_transform(National_stats, 4326)
Island <- National_stats %>%
  group_by(Nation, Date) %>%
  summarise(pop=sum(pop, na.rm = TRUE),
            Cases=sum(Cases, na.rm=TRUE),
            Death=sum(Death, na.rm=TRUE),
            Recovered=sum(Recovered, na.rm=TRUE))
Ireland_centroil = st_centroid(Island)

#extracting the longitude and latitude of the centroids to centre the circles at.
lng <- c(Ireland_centroil$geometry[[1]][1])
lat <- c(Ireland_centroil$geometry[[1]][2])

#creating map
Province_map <- leaflet() %>%
  setView(lng = lng, lat = lat,zoom = 6) %>% #setting the view to island of ireland centroid.
  addMapPane("background_map", zIndex = 410) %>%  # Level 1: bottom
  addMapPane("polygons", zIndex = 420) %>%        # Level 2: middle
  addMapPane("circles", zIndex = 430) %>%          # Level 3: top
  addProviderTiles(providers$Esri.WorldTopoMap,
                   options = pathOptions(pane = "background_map"))%>%
    addPolygons(data = Province_stats
                [Province_stats$Date == unique_date[3],],
                fillColor = ~provpalcases(Cases), 
                fillOpacity = 0.7,
                color = "#b2aeae", #boundary colour, need to use hex color codes.
                weight = 0.5, 
                smoothFactor = 0.2,
                label = lapply(Province_labs[3,] , htmltools::HTML),
                labelOptions = labeloptions,
                highlightOptions = highlight,
                options = pathOptions(pane = "polygons"),
                group = unique_date[3]) %>% #adding to group
  addLegend(pal = provpalcases,
              values = Province_stats$Cases[Province_stats$Date == unique_date[3]], 
              position = "topright", 
              title = "Number of Cases<br>",
              group = "county") %>%
  addPolygons(data = Province_stats
              [Province_stats$Date == unique_date[1],],
              fillColor = ~provpalcases(Cases), 
              color = "#b2aeae", # Need to use hex color codes.
              fillOpacity = 0.7, 
              weight = 0.5, 
              smoothFactor = 0.2,
              label = lapply(Province_labs[1,],
                htmltools::HTML),
              labelOptions = labeloptions,
              highlightOptions = highlight,
              options = pathOptions(pane = "polygons"),
              group = unique_date[1]) %>% #adding to group
    addPolygons(data = Province_stats
              [Province_stats$Date == unique_date[2],],
              fillColor = ~provpalcases(Cases), 
              color = "#b2aeae", # Need to use hex color codes.
              fillOpacity = 0.7, 
              weight = 0.5, 
              smoothFactor = 0.2,
              label = lapply(Province_labs[2,], htmltools::HTML),
              labelOptions = labeloptions,
              highlightOptions = highlight,
              options = pathOptions(pane = "polygons"),
              group = unique_date[2]) %>% #adding to group
    addCircles(data = Province_stats[Province_stats$Date == unique_date[3],],
               lng = provlng,
               lat = provlat,
               fillColor = ~provpalrate(rate),
               fillOpacity = 0.5,
               weight = 0.5, 
               color = "#FFFFFF",
               radius = 100000000*(Province_stats$rate)
               [Province_stats$Date == unique_date[3]],
               label = lapply(provcirlelabel, htmltools::HTML),
               labelOptions = labeloptions,
               highlightOptions = highlight,
               options = pathOptions(pane = "circles"),
               group = "Rate") %>% #adding to group
    addLegend(pal = provpalrate,
              values = Province_stats$rate[Province_stats$Date == unique_date[3]], 
              position = "bottomleft", 
              title = "Infection Rate<br>",
              group = "circles") %>%
  addLayersControl(overlayGroups = c("Rate",
                                     as.character(unique_date[1]),
                                     as.character(unique_date[2]),
                                     as.character(unique_date[3]))) #setting order of groups

# Display and save the map.
invisible(print(Province_map))
```